Summer 2024 Report

### 2024-12-12 Relations/Report.tex
Report updated, other tex files now redundant

        \item Does UN-lemma \emph{require} the decidability assumption?
      \end{itemize}
    \item META questions: Which of the ARS theorems can be made ``local/pointwise'',
    for example,
    \begin{itemize}
      \item NL+: $WCR(R) \land SN(x) \to CR(x)$?  (Yes?)
      \item NL++: $WCR(x) \land SN(x) \to CR(x)$? (No.)
    \end{itemize}
    \item Can "sequential" properties like RP, RP-, isWFseq, isWFseq-, $\omega$-bounded, etc.,
      be related to the cofinality property.
      \item What happens if properties involving sequences are reformulated for
    $R^r$? (Cannot decide whether the given sequence is finite or infinite.)
    \item Knaster-Tarski Lemma: Most general form? Alternative formulations?
    \item Knaster--Tarski in a well-founded setting: what additional
    hypotheses are necessary?
  \end{enumerate}

@@

\subsection{Main themes}
\begin{enumerate}
  \item Clean up the implications around recurrence properties:
  \begin{itemize}
    \item RP iff cofinality and Boundedness?
    \item What is the ``recurrent'' generalization of WN?
    \item What is the ``recurrent'' generalization of SN?
    \item The role of WCR in the proof of our generalization of Thm. 1.2.3(iii)
  \end{itemize}
  \item Complete the ``design'' of the chart(s) with implications?
  \item Complete the proof of Theorem 1.2.3
  \item Address outstanding questions about wellfoundness implications.
  \begin{itemize}
    \item WFseq implies WFmin? (Assuming decMin)
    \item WFacc implies $WFmin_0$? (Assuming decMin)
    \item DecMin suffices for $SN\to WN$, but what is the optimal \emph{necessary} condition?
    \item Relationship between these, and De Morgan's laws.
  \end{itemize}
\end{enumerate}

\subsection{General todo list}
\begin{enumerate}
  \item Look at Exercise 1.3.22
  \item  Make the first argument to "being a normal form" implicit?
  \item Define WNFP to be:
  \[\forall a b c \to b \in NF \to a R^* b \to a R^* c \to c R^* b \]
  \item Does WNFP and WN imply CR?  Looks like it!
  \item  Another interesting one? $WN^*$: hereditarily $WN$ ($WN$ for the reduction graph starting from a given element $x$.)

  Could even consider $SN^*$, $CR^*$ etc.
  \item Define: Inc
  \item Define: Ind
  \item Define: $\mathrm{CR}^{\le 1}$
  \item capture all classical properties/asssumptions we might want to use and justify their use.
  \item Principles we are considering
  \begin{itemize}
    \item Decidability of equality on $A$
    \item Decidability of the relation $R$
    \item When relevant, decidability of the given precicate $P$
    \item Not-not-closure of being accessible
    \item Not-not-closure of $R$
    \item Not-not-closure of $P$
    \item Decidability of whether a given element is an $R$-normal form
    \item When relevant, local/element-wise versions of the above
    \item Markov's principle:
    \begin{itemize}
      \item In $wfMin_0 \to wfSeq$.

    \end{itemize}
  \end{itemize}
  \item Write a guide to the code
  \item More laws about closure operations: Monotonicity, idempotency
\end{enumerate}

\subsection*{Week starting 11th November}
\begin{enumerate}
  \item[\cmark] In the report, give a classical proof that the $Inc$ Property
  implies that no $R$-increasing sequence is bounded.

  Conclude that every relation satisfying $Inc$ also satisfies $RP$.

  Therefore, the proof of Theorem 1.2.3(iii) using $RP$ instead of $Inc$
  is generalizing the claim of the theorem (as stated in [TeReSe]).
  \item Complete the proof in ARS.agda of Theorem 1.2.3.(iii)
  \item Organize notes.tex into three sections: TODO, Open Qs, What we've proved. Make sure all notes from the other
  tex files are sorted here correctly.
  \item Tidy up and standardize the notation used in ars.agda and other agda files
  \item Carry on working through report.tex.
\end{enumerate}

@@

\subsection{Notes from 08.29}
\begin{enumerate}
  \item Try to ``fix'' CP by changing $R$ into $R^r$
  \item Finish formulating the ``compactness property".  Two candidates:
  \begin{itemize}
    \item ``Every cocone for a given infinite sequence loops back to some point in the sequence"
    \item ``If an $R^*$-sequence has a cocone then it's constant after some point on.''
  \end{itemize}
  \item Add the definition of "recurrent element"
\end{enumerate}

\subsection{09.05}
\begin{enumerate}
  \item done: WN, UN, RP imply SN (classical)
  \item need: omega bounded and RP implies SN (classical)
  \item need: omega bounded and RP implies SN (constructive)
  \item do we need $s_0 \to^* s_c$ in the proof?
  \item is $\omega$-bounded implied by WN and the following confluence property:
  $a \to^* n \in NF$, and $a \to b$ then $b \to^* n $.
  \item Remark before 1.2? (Exercise 1.3.21)
  \item Make the first argument to "being a normal form" implicit?
\end{enumerate}

\newcommand{\RP}{\mathsf{RP}}
\newcommand{\NF}{\mathsf{NF}}
\newcommand{\UN}{\mathsf{UN}}
\newcommand{\UNto}{\mathsf{UN}{\to}}
\newcommand{\SN}{\mathsf{SN}}
\newcommand{\WN}{\mathsf{WN}}
\newcommand{\from}{\leftarrow}

\subsection*{Additional remarks about Theorem 1.2.3.}
\begin{itemize}
  \item It seems that part (i) of the theorem can be strengthened to only
  assume $\UNto(R)$ instead of $\UN(R)$:
  \[ \WN(R) \land \UNto(R) \to \omega{-}\text{bounded}(R) \tag{i} \]
  \textcolor{red}{(See ARS.agda i+)}
  \item The analogue of part (ii) of the theorem is unprovable, namely
  \[ \omega{-}\text{bounded}(R) \land \RP(R) \to \SN(R) \tag{ii}\]
  For a countermodel, take the one-element loop $a \to a$.

  (It is both $\omega${-}bounded and satisfies RP.)

  Our hypothesis RP is indeed weaker than ``Inc'' from Terese.
  \item The goal \texttt{iii-lemma :  $WN(R) \to WCR(R) \to \omega{-}\text{bounded}(R)$}
  seems either very difficult or impossible.

@@

\end{itemize}
\subsection{09.10 (midweek updata)}
Is it possible to weaken RP? A possible suggestion:

$RP- :: \omega-bounded \to  \exists i. b \to^* f i$

Does:

$RP \to RP-$

Does:

$RP- \land \; WCR \to RP$

(what other property with $RP-$ would give us $RP$?).

>>>>

        \item Does UN-lemma \emph{require} the decidability assumption?
      \end{itemize}
    \item META questions: Which of the ARS theorems can be made ``local/pointwise'',
    for example,
    \begin{itemize}
      \item NL+: $WCR(R) \land SN(x) \to CR(x)$?  (Yes?)
      \item NL++: $WCR(x) \land SN(x) \to CR(x)$? (No.)
    \end{itemize}
    \item Can "sequential" properties like RP, RP-, isWFseq, isWFseq-, $\omega$-bounded, etc.,
      be related to the cofinality property.
      \item What happens if properties involving sequences are reformulated for
    $R^r$? (Cannot decide whether the given sequence is finite or infinite.)
    \item Knaster-Tarski Lemma: Most general form? Alternative formulations?
    \item Knaster--Tarski in a well-founded setting: what additional
    hypotheses are necessary?
    \item Show SN and WNFP $\to$ CR : SN R $to$ WNFP R $to$ CR R (or, what combination of properties with WNFP gets us to CR)
    \item Major: Find a classical or constructive solution to theorem 1-2-3 ii. Note, we have a counterexample showing that we 
    cannot solve this whilst using recurrence property and boundedness rather than ind(uctive) and inc(reasing). The counterexample is the self loop on an element. 
    \item is $\omega$-bounded implied by WN and the following confluence property:
    $a \to^* n \in NF$, and $a \to b$ then $b \to^* n $. (This is a weaker form of WNFP)
  \end{enumerate}

@@

\subsection{Main themes}
\begin{enumerate}
  \item Clean up the implications around recurrence properties:
  \begin{itemize}
    \item RP iff cofinality and Boundedness?
    \item What is the ``recurrent'' generalization of WN?
    \item What is the ``recurrent'' generalization of SN?
    \item The role of WCR in the proof of our generalization of Thm. 1.2.3(iii)
  \end{itemize}
  \item Complete the ``design'' of the chart(s) with implications?
  \begin{itemize}
    \item   Look at Exercise 1.3.22 (Show UN $\to$ UN$^\to$ but not conversely. Show a counterexample to WCR and UN$^\to$ $\to$ UN. Find a counter example to WCR and WN $\to$ Ind)
  \end{itemize}
  \item Complete the proof of Theorem 1.2.3
  \item Address outstanding questions about wellfoundness implications.
  \begin{itemize}
    \item WFseq implies WFmin? (Assuming decMin)
    \item WFacc implies $WFmin_0$? (Assuming decMin)
    \item DecMin suffices for $SN\to WN$, but what is the optimal \emph{necessary} condition?
    \item Relationship between these, and De Morgan's laws.
  \end{itemize}
  \item Complete the classical properties section
  \item Complete the guide to the code section
  \item Miscellaneous
  \begin{itemize}
    \item   Make the first argument to "being a normal form" implicit?
    \item   Define: Inc in report
    \item   Define: Ind in report 
    \item   Define: $\mathrm{CR}^{\le 1}$ in report
    \item   More laws about closure operations: Monotonicity, idempotency
    \item   Tidy up and standardize the notation used in ars.agda and other agda files

  \end{itemize}
\end{enumerate}

\subsection{General todo list}
\begin{enumerate}
  \item  Another interesting one? $WN^*$: hereditarily $WN$ ($WN$ for the reduction graph starting from a given element $x$). Could even consider $SN^*$, $CR^*$ etc. \sacomment{Is this what we have already shown 
  in the reduction closure properties? If so we can delete this. If not it should be moved to open questions}
\end{enumerate}

@@

\subsection{Notes from 08.29}
\begin{enumerate}
  \item Finish formulating the ``compactness property".  Two candidates:
  \begin{itemize}
    \item ``Every cocone for a given infinite sequence loops back to some point in the sequence"
    \item ``If an $R^*$-sequence has a cocone then it's constant after some point on.''
  \end{itemize}
\end{enumerate} \sacomment{This is our recurrence property right? If so, we can just delete this.}

\subsection*{Additional remarks about Theorem 1.2.3.}
\begin{itemize}

  \item The goal \texttt{iii-lemma :  $WN(R) \to WCR(R) \to \omega{-}\text{bounded}(R)$}
  seems either very difficult or impossible.

@@
